### 2016-08-25 section/1_under.tex
finish sec 2

Previous work has identified a wide variety of performance root-cause 
categories. However, existing taxonomies do not satisfy all the three 
requirements below and hence cannot be directly used to guide our 
diagnosis-tool
design.
\begin{enumerate}
\item Coverage: covering
a big portion of real-world inefficient loop problems; 
\item Actionability: each root-cause category should be informative enough 
to help developers
decide how to fix a performance problem; 
\item Generality: application-specific root causes will not work, as we
hope to build diagnosis tools that can automatically identify root causes 
without developers' help.
\end{enumerate}

We divide all inefficient loops into two main root-cause categories,
\textit{resultless} and \textit{redundancy}, and many sub-categories under
these two.
The details of these categories will be presented below.

Our discussion below will qualitatively show that 
our taxonomy is general, not application-specific, and
cover common loop inefficiency problems.
We will quantitatively measure the coverage, actionability, and 
generality of our taxonomy
using real-world performance-bug suite in Section \ref{sec:eval_taxonomy}.

\subsection{Resultless loops}
\label{sec:study_resultless}
Resultless loops spend a lot of time in computation that does not
produce results useful after the loop (i.e., computation without side effects).
They can be further categorized to four sub-types 
based on which part of the loop is
(not) producing useful results. We explain these four sub-types below, 
as well as
how they can be fixed.

{\textbf{0*}}: 
This type of loops never produce any results in any iteration.
They are rare in mature software systems.

{\textit{How to fix?}} This type of loops should simply be deleted from the program.
%Question: dynamic or static?

@@

{\textbf{0*1?}}:
This type of loops only produce results in the last iteration, if any. 
They are often related to search: check a sequence of elements one
by one until the right one is found.
Clearly, whether these loops are efficient or not depends on the
workload. When indeed inefficient, they are often
fixed by data-structure changes.
An example is shown in Figure \ref{fig:Mozilla347306}.
Large JavaScript files often fill the \texttt{script} list with 
tens of thousands of nodes and cause poor performance.

{\textit{How to fix?}}
To make a linear search more efficient, it usually requires a data
structure change. For example, 
the patch for the bug in Figure \ref{fig:Mozilla347306}
simply replaced the \texttt{script} list with a hash table.
% This improve the performance by 52.60{\bf X}.

@@

{\textbf{[0$|$1]*}}:
Each iteration in this type of loops may or may not produce results.
For some workload, the majority of iterations do not produce
results and cause performance problems perceived by users.
%These problems are often solved
%by adding a simple check to conditionally skip the expensive loop.
Figure~\ref{fig:GCC46401} shows such an example.
%how users perceive this problem
%explain how every iteration may or may not produce any results
%explain the fix
Users complained that compilation became extremely slow when
the \texttt{-Wsequence-point} checking is enabled.
The slowness was caused
by the \texttt{for} loop in the figure. As the algorithm
behind this loop has quadratic complexity in the number of operands in an expression, 
programs with long expressions suffer severe slow-downs.
After further diagnosis, developers observed that this loop rarely
had any side-effects, as the \texttt{if} condition was rarely satisfied.
At the end, the patch greatly improved performance through a
simple checking that allows the loop to be
skipped in most cases.

@@

{\textbf{1*}}:
Loops in this category always generate results in almost all iterations. 
They are inefficient because their results are useless due to
some high-level semantic reasons.
Understanding and fixing this type of inefficiency problems often require
deep understanding of the program and are difficult to automate.
For example, several Mozilla performance problems are caused by 
loops that contain intensive GUI operations whose graphical outcome may not
be observed by humans and hence can be optimized.

{\textit{How to fix?}}
Since a deep understanding of software semantics is required
to understand the inefficiency of these loops,
the fix strategies for these loops will likely vary from case to case.
%At high level, the patch often involves skip some computation that produces
%side effects that are useless due to semantic reasons.

@@

{\textbf{Cross-iteration Redundancy}}:
Loop iteration is the redundancy unit here:
one iteration repeats
what was already done by an earlier iteration of the same loop.
Here, we consider a recursive function as a loop, treating one function-call
instance as one loop iteration.

{\textit{How to fix?}}
Intuitively, most redundancy problems can be fixed through memoization or
batching --- either caching the earlier computation results and skip
some of the following iterations; or combining multiple iterations' work
together.
For example, the GCC bug shown in Figure~\ref{fig:GCC27733} is caused by
redundant computation across different invocations of recursive function
\texttt{mult\_alg}. The patch essentially enables memoization.
%Another bug similar to GCC\#27733 is shown in Figure~\ref{fig:GCC1687}. 
%This bug is fixed by using a hash table to avoid visiting the same sub-tree twice.
\comment{
Mozilla\#490742 in Figure~\ref{fig:Mozilla490742} represents a slightly different type of cross-iteration
redundancy. The inefficient loop in this case saves one URL into the ``Favorite
Links'' database in each iteration. One database transaction in each iteration
turns out to be to time consuming, with too much redundant work across
iterations. At the end, developers decide to batch all database updates into 
one big transaction, which speeds up some workload like 
bookmarking 50 tabs 
from popping up timeout windows to not blocking.
%xxx from several minutes
%to a couple of seconds.
}

{\textbf{Cross-loop Redundancy}}:
A whole loop is the redundancy unit:
one dynamic instance of a loop spends a big chunk, if not all, of its
computation in repeating the work already done by an
earlier instance of the same loop.

{\textit{How to fix?}}
Just like that in cross-iteration redundancy, memoization and batching
are the typical fix strategies for this type of redundancy problems ---
caching the earlier computation results and skip following redundant loops;
or combining multiple loop instances together to avoid or alleviate redundancy.
Mozilla\#477564 shown in Figure~\ref{fig:Mozilla477564} is an example for this type of bugs. 
The buggy loop counts how many previous siblings of the input \texttt{aNode} have the same name and URI. 
There is an outer loop, not shown in the figure, that repeatedly updates
\texttt{aNode} to be its next sibling and calls
\texttt{sss\_xph\_generate} with the new \texttt{aNode}. 
This bug is fixed by adding an extra field for each node to save the calculated 
count, so that a new count value can be calculated by simply adding one to the
saved count value of the nearest previous sibling with the same name and URI.

@@

\subsection{Discussion}
Coverage, actionability, and generality are the principles behind our taxonomy
design. Intuitively, the categories above cover a lot of common inefficient
loop problems; each sub-category is concrete enough to guide the design of a
failure diagnosis tool; none of the categories above involve any application
specific knowledges or heuristics. We will further assess our taxonomy
through the design of \Tool (Section \ref{sec:design}) and 
a real-world bug study (Section \ref{sec:eval_taxonomy}).

Of course, our taxonomy does not cover \emph{all} loop inefficiency problems.
For example, some loops may be vulnerable to false sharing problems or lock
contention problems, which are out of the scope of our taxonomy.

>>>>

To guide the design of automated diagnosis tools, we need a
taxonomy for inefficient loops that satisfies three requirements:
(1) \textit{Coverage}, covering
a big portion of real-world inefficient loop problems; 
(2) \textit{Actionability}, each root-cause category being informative enough 
to help developers fix a performance problem; 
(3) \textit{Generality}, allowing automated diagnosis to be applied to many
applications.

Previous work identified a wide variety of performance root-cause 
categories. However, existing taxonomies do not satisfy all these 
requirements and hence cannot be directly used by us.
Therefore, we have designed our own taxonomy,
as presented below.
We will discuss how our taxonomy meets these three requirements
qualitatively in this section and quantitatively 
using real-world performance-bug suite in Section \ref{sec:eval_taxonomy}.

\subsection{Resultless loops}
\label{sec:study_resultless}
Resultless loops spend a lot of time in computation that does not
produce results useful after the loop (i.e., no side effects).
They can be further categorized to four sub-types 
based on which part of the loop is
(not) producing useful results.

{\textbf{0*}} loops
never produce any results in any iteration.
They are rare in mature software systems.

{\textit{How to fix?}} They should be deleted from software.
%Question: dynamic or static?

@@

{\textbf{0*1?}} loops only produce results in the last iteration, if any. 
They are often related to search: check a sequence of elements one
by one until the right one is found.
Whether these loops are efficient or not depends on the
workload. 
An example is shown in Figure \ref{fig:Mozilla347306}.
Large JavaScript files often fill the \texttt{script} list with 
tens of thousands of nodes and cause poor performance.

{\textit{How to fix?}}
They are often fixed by data-structure changes.
For example, the patch for the bug in Figure \ref{fig:Mozilla347306}
simply replaced the \texttt{script} list with a hash table.
% This improve the performance by 52.60{\bf X}.

@@

{\textbf{[0$|$1]*}} loops may or may not produce results in each
iteration.
For some workload, the majority of iterations do not produce
results and cause performance problems perceived by users.
%These problems are often solved
%by adding a simple check to conditionally skip the expensive loop.
Figure~\ref{fig:GCC46401} shows such an example.
%how users perceive this problem
%explain how every iteration may or may not produce any results
%explain the fix
Users complained that compilation became extremely slow when
the \texttt{-Wsequence-point} checking is enabled.
The slowness was caused
by the \texttt{for} loop in the figure. As the algorithm
behind this loop has quadratic complexity in the number of operands in an expression, 
programs with long expressions suffer severe slow-downs.
After further diagnosis, developers observed that this loop rarely
had any side-effects, as the \texttt{if} condition was rarely satisfied.
%At the end, the patch greatly improved performance through a
%simple checking that allows the loop to be skipped in most cases.

@@

{\textbf{1*}} loops in this category always generate results in almost all iterations. 
They are inefficient because their results are useless due to high-level semantic reasons.
%Understanding and fixing this type of inefficiency problems often require
%deep understanding of the program and are difficult to automate.
For example, several Mozilla performance problems are caused by 
loops that contain intensive GUI operations whose graphical outcome may not
be observed by humans and hence can be optimized.

{\textit{How to fix?}}
Since a deep understanding of software semantics is required
to understand the inefficiency of these loops,
the fix strategies for these loops likely vary from case to case
and are difficult to automate.
%At high level, the patch often involves skip some computation that produces
%side effects that are useless due to semantic reasons.

@@

{\textbf{Cross-iteration Redundancy}}:
one iteration repeats
what was already done by an earlier iteration of the same loop.
Here, we consider a recursive function as a loop, treating one function-call
instance as one loop iteration.

{\textit{How to fix?}}
Intuitively, most redundancy problems can be fixed through memoization or
batching --- either caching the earlier computation results and skip
some following iterations; or combining multiple iterations' work
together.
For example, the bug shown in Figure~\ref{fig:GCC27733} is caused by
redundant computation across different invocations of recursive function
\texttt{mult\_alg}. The patch essentially enables memoization.
%Another bug similar to GCC\#27733 is shown in Figure~\ref{fig:GCC1687}. 
%This bug is fixed by using a hash table to avoid visiting the same sub-tree twice.
\comment{
Mozilla\#490742 in Figure~\ref{fig:Mozilla490742} represents a slightly different type of cross-iteration
redundancy. The inefficient loop in this case saves one URL into the ``Favorite
Links'' database in each iteration. One database transaction in each iteration
turns out to be to time consuming, with too much redundant work across
iterations. At the end, developers decide to batch all database updates into 
one big transaction, which speeds up some workload like 
bookmarking 50 tabs 
from popping up timeout windows to not blocking.
%xxx from several minutes
%to a couple of seconds.
}

{\textbf{Cross-loop Redundancy}}:
one dynamic instance of a loop spends a big chunk, if not all, of its
computation in repeating the work already done by an
earlier instance of the same loop.

{\textit{How to fix?}}
Just like that in cross-iteration redundancy, memoization and batching
are the typical fix strategies for this type of loops.
Mozilla\#477564 shown in Figure~\ref{fig:Mozilla477564} is an example. 
The buggy loop counts how many previous siblings of the input \texttt{aNode} have the same name and URI. 
There is an outer loop, not shown in the figure, that repeatedly updates
\texttt{aNode} to be its next sibling and calls
\texttt{sss\_xph\_generate} with the new \texttt{aNode}. 
This bug is fixed by adding an extra field for each node to save the calculated 
count, so that a new count value can be calculated by simply adding one to the
saved count value of the nearest previous sibling with the same name and URI.

@@

\noindent\textbf{Threats to Validity}
Intuitively, the categories above cover a lot of common inefficient
loop problems; each sub-category is concrete enough to guide the design of a
failure diagnosis tool; none of the categories above involve any application
specific knowledges or heuristics.
Of course, our taxonomy does not cover \emph{all} loop inefficiency problems.
For example, some loops may be vulnerable to false sharing problems or lock
contention problems, which are out of the scope of our taxonomy.
